---
title: "Análise dos dados em Pimentel-Gomes (2009)"
author: "Walmes Zeviani"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Análise dos dados em Pimentel-Gomes (2009)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Essa *vignette* tem como objetivo fazer a análise dos conjuntos de dados
do livro Pimentel-Gomes (2009), a décima quinta edição da obra **Curso
de Estatística Experimental**.

Para instalar o [`legTools`][], faça

```{r, eval=FALSE}
library(devtools)
install_git("http://git.leg.ufpr.br/leg/legTools.git")

```

```{r}
library(legTools)
packageVersion("legTools")

```
[`legTools`]: http://git.leg.ufpr.br/leg/legTools


## Delineamento Inteiramente Casualizado

Considere os dados disponíveis na página 62 referentes à valores
fictícios de um experimento de alimentação de porcos no qual se usou 4
níveis do fator ração. O experimento foi instalado em um delineamento
inteiramente casualizado com 5 animais (unidades experimentais) para
cada nível de ração. A variável resposta medida foi o ganho de peso dos
animais (kg).

Esses dados fazem parte do pacote `legTools` e estão no *data frame*
`wgpigs`. Alternativamente, pode-se fazer o download do mesmo no formato
texto separado por tabulação.

```{r}
## url <- "http://blog.leg.ufpr.br/~leg/legTools/dataset/wgpigs.txt"
## browseURL(url)
## wgpigs <- read.table(file=url, header=TRUE, sep="\t")

library(legTools)

data(wgpigs)
str(wgpigs)

```



<!--==================================================================== -->

Vignettes are long form documentation commonly included in
packages. Because they are part of the distribution of the package, they
need to be as compact as possible. The `html_vignette` output type
provides a custom style sheet (and tweaks some options) to ensure that
the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap
  style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata
block above. These are required in order to instruct R how to build the
vignette. Note that you should change the `title` field and the
`\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override
this theme you can specify your own CSS in the document metadata as
follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two
images side-by-side.

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in
**knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$,
footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
